---
title: "EDLD 652 Homework"
author: "Eliott Doyle, Diana DeWald"
date: "2/8/2022"
output:
  html_document: default
  pdf_document: default
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggrepel)
library(gghighlight)
library(ggtext)
library(ggforce)
library(geomtextpath)
transit_cost <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-01-05/transit_cost.csv')
```

## Question 1: 

### Use the transit costs data to reproduce the following plot. To do so, you will need to do a small amount of data cleaning, then calculate the means and standard errors (of the mean) for each country. Please filter to only counties with at least three observations. To use actual country names, rather than abbreviations, join your dataset with the output from the following

# removing the last 7 rows which contain many NA values

```{r q1, echo = FALSE, warning = FALSE, message = FALSE}
# removing the last 7 rows which contain many NA values
transit_cost <- transit_cost[-c(538, 539, 540, 541, 542, 543, 544), ]

library(countrycode)
country_codes <- countrycode::codelist %>% 
  select(country_name = country.name.en, country = ecb)

# joining datasets by country codes and removing other NAs

transit_new <- left_join(transit_cost, country_codes, by = "country")
transit_new <- transit_new[-c(141, 481, 482), ]
transit_new <-  subset(transit_new, select = -c(reference, end_year) )

# removing countries with less than 3 observations
transit_new <- transit_new %>%
  group_by(country_name) %>% 
  filter(n() >= 3)
transit_new$real_cost <- as.numeric(transit_new$real_cost)
transit_new$country_name <- as.factor(transit_new$country_name)

# adding mean and standard deviation columns
transit_new.1 <- transit_new %>%
  group_by(country_name) %>%
  summarize(mean_realcost = mean(real_cost, na.rm = TRUE),
            sd_realcost = sd(real_cost, na.rm = TRUE))
  
# adding standard error column
transit_new <- transit_new %>%
  group_by(country_name) %>%
  summarize(mean_realcost = mean(real_cost, na.rm = TRUE),
            sd_realcost = sd(real_cost, na.rm = TRUE))
  transit_new.1 <- transit_new %>%
    group_by(country_name) %>%
    mutate(sem_realcost = mean(sd_realcost)/sqrt(length(transit_new[])))
  
  
# preparing the plot
  
x <- c(10000, 20000, 30000)
cost <- data.frame(x)

se.1 <- (transit_new.1$mean_realcost - transit_new.1$sem_realcost)
se.2 <- (transit_new.1$mean_realcost + transit_new.1$sem_realcost)

transit_1 <- transit_new.1 %>% 
  ggplot(aes(x = mean_realcost,
             y = reorder(country_name, mean_realcost))) +
  geom_point(color = "cornflowerblue", size = 2) +
  xlim(0, 30000) +
  theme_classic() +
  labs(
    title = "Cost to build transit systems vary across countries",
    x = "Real Cost (In millions of dollars)",
    y = "Country",
    caption = "Data provided through #tidytuesday by the Transit Costs Project") +
    geom_linerange(aes(xmin= se.1, xmax= se.2), color = "gray40", size = 0.5) +
    theme(plot.title = element_text(hjust = -0.9, vjust=2.12)) + 
    theme(plot.caption=element_text(size=10, hjust= 0.5)) +
    geom_vline(data = cost, aes(xintercept = x), color = "darkgrey")

transit_1  
```


## Question 2: 

### Visualize the same relation, but displaying the uncertainty using an alternative method of your choosing.

```{r q2, echo = FALSE, warning = FALSE, message = FALSE}
# 95%, 90%, and 80% to visualize uncertainty

transit_2 <- transit_new.1 %>% 
  ggplot(aes(x = mean_realcost,
             y = reorder(country_name, mean_realcost))) +
  geom_point(color = "gray63", size = 2) +
  xlim(0, 30000) +
  theme_classic() +
  labs(
    title = "Cost to build transit systems vary across countries (Pt. 2)",
    x = "Real Cost (In millions of dollars)",
    y = "Country",
    caption = "Horizontal lines represent 95%, 90%, and 80% CI") +
    geom_vline(data = cost, aes(xintercept = x), color = "darkgrey") +
  geom_errorbar(aes(xmin = mean_realcost + qnorm(.025)*sem_realcost, 
                    xmax = mean_realcost + qnorm(.975)*sem_realcost),
                color = "steelblue1",
                width = 0.2,
                size = 0.3) + # 95% CI 
   geom_errorbar(aes(xmin = mean_realcost + qnorm(.05)*sem_realcost, 
                    xmax = mean_realcost + qnorm(.95)*sem_realcost),
               color = "steelblue3",
                width = 0.2,
                size = 0.6) +  # 90% CI
  geom_errorbar(aes(xmin = mean_realcost + qnorm(.1)*sem_realcost, 
                    xmax = mean_realcost + qnorm(.9)*sem_realcost),
                color = "steelblue4",
                width = 0.2,
                size = 0.8)  # 80% 
    
    
transit_2
```


## Question 3: 

### Compute the mean length and real_cost by city. Reproduce the following plot.

#### Hint: Look into scale_size_binned()

```{r q3, echo = FALSE, warning = FALSE, message = FALSE}
#View(transit_new)
#View(transit_cost)
#mean length and realcost by city
q3 <- transit_cost %>% 
  select(country, city, length, real_cost) %>% 
  group_by(country, city)
q3$real_cost <- as.numeric(q3$real_cost)
#View(q3)
#head(q3)
q3_calc <- q3 %>% 
  summarize(mean_length = mean(length, na.rm = TRUE),
         mean_real_cost = mean(real_cost, na.rm = TRUE),
         count=n()) 
#View(q3_calc)
q3_fig <- q3_calc %>% 
  ggplot(aes(x = mean_length,
             y = mean_real_cost
             )) +
  geom_point(aes(size = count), color = "#b452cd") +
  scale_size_binned() +
  gghighlight(country == "IN") +
  geom_text_repel(data = dplyr::filter(q3_calc,
                                country == "IN"),
                  aes(label = city),
                  min.segment.length = 0,
                  box.padding = 0.2) +
  scale_x_log10() +
  scale_y_log10(labels = scales::dollar_format()) +
  theme_minimal(base_size = 14) +
  labs(title = "Longer transit systems tend to cost more",
       subtitle = "<span style = 'color: #b452cd'>India</span> has among the most transit systems in the world",
       x = "Average length",
       y = "Average cost",
       caption = "Note the log transformations to the axes",
       size = "Number of transit systems") +
  theme(plot.title.position = "plot", #to left-align title
        plot.caption.position =  "plot", # & subtitle
        legend.position="bottom",
        plot.subtitle = element_markdown()) #needed to get subtitle color to work
q3_fig
```


## Question 4:

### Using basically the same data, reproduce the following plot. Note you’ll need the country_name column in your dataset.


```{r q4.2, echo = FALSE, warning = FALSE, message = FALSE}
# new df, joining by country codes
q4_calc <- q3_calc
q4_calc <- left_join(q4_calc, country_codes, by = "country")
#View(q4_calc)
q4_fig <- q4_calc %>% 
  ggplot(aes(x = mean_length,
             y = mean_real_cost
             )) +
  geom_point(aes(size = count), color = "#5C98AD") +
  geom_mark_ellipse(aes(group = country_name, label = country_name),
                    color = "#F7B2C1",
                    data = filter(drop_na(q4_calc),
                                  country_name == "United States"),
                    size = .5,
                    con.colour = "#F7B2C1",
                    con.size = .3,
                    con.type = "straight",
                    label.colour = "#5C98AD",
                    show.legend = FALSE) +
  geom_label_repel(data = dplyr::filter(q4_calc,
                                country == "US"),
    aes(label = city),
    box.padding = 0.35,
    min.segment.length = 0,
    point.padding = 0.5
  ) +
  guides(color = "none") +
  scale_size_binned() +
  gghighlight(country_name == "United States") +
  scale_x_log10() +
  scale_y_log10(labels = scales::dollar_format()) +
  theme_minimal(base_size = 14) +
  labs(title = "Longer transit systems tend to cost more",
       x = "Average length",
       y = "Average cost",
       caption = "Note the log transformations to the axes",
       size = "Number of transit systems") +
  theme(plot.title.position = "plot",
        plot.caption.position =  "plot",
        legend.position="bottom",
        plot.subtitle = element_markdown()) 
q4_fig
```


## Question 5:

### Use the crime dataset to run the following code and fit the corresponding model. Note, it may take a moment to run.

```{r q5setup, include = FALSE}
#reading the data in as a .csv file, converting & saving .parquet file. the .csv file has been deleted from the folder so that code won't work anymore
#crime <- read_csv("crime.csv")
library(arrow)
#write_parquet(crime, here::here("crime.parquet"))
crime <- read_parquet(here::here("crime.parquet"))
names(crime) <- tolower(names(crime))
#View(crime)
```


```{r q5.1, echo = FALSE, warning = FALSE, message = FALSE}
 model_data <- crime %>% 
  mutate(neighborhood_id = relevel(factor(neighborhood_id), ref = "barnum"))
 m <- glm(is_crime ~ neighborhood_id, 
         data = model_data,
         family = "binomial")
#This model uses neighborhood to predict whether a crime occurred or not. The reference group has been set to the “barnum” neighborhood, and the coefficents are all in comparison to this neighborhood. Extract the output using broom::tidy
 tidied <- broom::tidy(m)
# Divide the probability space, [0, 1], into even bins of your choosing. For example, for 20 bins I could run the following
 ppoints(20)
 #The coefficients (tidied$estimate) for each district in the model represent the difference in crime rates between the corresponding neighborhood the Barnum neighborhood. These are reported on a log scale and can be exponentiated to provide the odds. For example the Athmar-Park neighborhood was estimated as 1.13 times more likely to have a crime occur than the Barnum neighborhood. This is the point estimate, which is our “best guess” as to the true difference, and the likelihood of alternative differences are distributed around this point with a standard deviation equal to the standard error. We can simulate data from this distribution, if we choose, or instead just use the distribution to calculate different quantiles.
 
 #The qnorm function transforms probabilities, such as those we generated with ppoints, into values according to some pre-defined normal distribution (by default it is a standard normal with a mean of zero and standard deviation of 1). For example qnorm(.75, mean = 100, sd = 10) provides the 75th percentile value from a distribution with a mean of 100 and a standard deviation of 10. We can therefore use qnorm in conjunction with ppoints to better understand the sampling distribution and, ultimately, communicate uncertainty. For example the following code generates the values corresponding to ppoints(20), or 2.5th to 97.5th percentiles of the distributions in 5 percentile “jumps”, for the difference in crime rates on the log scale for Barnum and Regis neighborhoods.
 regis <- tidied %>% 
  filter(term == "neighborhood_idregis")
 qnorm(ppoints(20), 
      mean = regis$estimate,
      sd = regis$std.error)
```

#### The following plot displays a discretized version of the probability space for the differences in crime between the neighborhoods. Replicate this plot, but comparing the Barnum neighborhood to the Barnum-West neighborhood. Make sure to put the values in a data frame, and create a new variable stating whether the difference is greater than zero (which you will use to fill by). Note that you do not need to replicate the colors in the subtitle to match the balls, as I have, but if you’d like to you should likely use the ggtext package.

#### Note: Your probabilities will not directly correspond with the p values, which are essentially twice the probability you are displaying (because the test is two-tailed).

```{r q5.2, echo = FALSE, warning = FALSE, message = FALSE}
barnumwest <- tidied %>% 
  filter(term == "neighborhood_idbarnum-west")
head(barnumwest)
qnorm(ppoints(20),
      barnumwest$estimate,
      barnumwest$std.error)
barnum <- tidied %>% 
  filter(term == "neighborhood_idbarnum")
qnorm(ppoints(20),
      barnum$estimate,
      barnum$std.error)

discretized <- data.frame(
 x = qnorm(ppoints(50), 1.02, 0.9)
 ) %>% 
 mutate(winner = ifelse(x > 0, "#b1daf4", "#f8f1a9"))

# head(discretized)


ggplot(discretized, aes(x)) +
  geom_dotplot(aes(fill = winner), binwidth = 0.26) +
  geom_vline(xintercept = 0, 
             color = "red", 
             linetype = 1, 
             size = 1) +
  scale_fill_identity(guide = "none") +
  scale_y_continuous(name = "", 
                     breaks = NULL) +
  theme_classic()

# qnorm(ppoints(20), 
#      mean = regis$estimate,
#      sd = regis$std.error)

```